Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2021-04-29 paper/ifo-introduction.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance. 
These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% CW 
To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}. 
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
%Scorpius X-1 is an LMXB that is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.

% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'' (using light to capture sound) and present a range of example analysis techniques for educators to use. 
% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
%% I don't like the use of ``freedom to explore''. 
% removed Additionally - I don't think it is neccessary - hannah
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}. 
% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world. 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
%\han{look up Sarah Farmer's project}

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyze the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
% removed as there's not space to really say what this means here: to varying degrees of success.
This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory. 
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
We collect the software and scripts used to produce this work in Appendix~\ref{app:code}.

>>>>

In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
Gravitational waves are a prediction of Albert Einstein's theory of General Relativity~\cite{Einstein:1916}, 
They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance. 
These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% CW 
To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}. 
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
The rotation frequency of the neutron star in an LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
%Scorpius X-1 is an LMXB that is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.

% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'', using light to capture sound, and present a range of example analysis techniques for educators to use. 
% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
%% I don't like the use of ``freedom to explore''. 
% removed Additionally - I don't think it is neccessary - hannah
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}. 
% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world. 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
%\han{look up Sarah Farmer's project}

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we observe a wandering frequency signal and analyze it using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
% removed as there's not space to really say what this means here: to varying degrees of success.
This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory. 
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
We present the software and scripts used to produce this work in Appendix~\ref{app:code}.

### 2021-04-29 paper/ifo-supplementary-material.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

In Section~V in the main article, we use the interferometer as an optical microphone; audio signals are used to vibrate the interferometer mirrors, the changing interference pattern is recorded, and we aim to recover the original signal from the resulting timeseries data.
In this Supplementary Material we include a selection of signal processing techniques applied to the data which may be considered for undergraduate laboratory examples. 
We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.

@@

\begin{figure}
	\begin{center}
    %% trim order is <left> <lower> <right> <upper>
    % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
	\includegraphics[width=.5\textwidth]
% trim={1cm 3cm 1cm 1cm}]
{figures/combined_results_full_melatos_labelled.pdf}
%notch_and_wiener_superplot_v2.pdf}
	\end{center}
	\caption{\label{fig:op_mic_results}
	%\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
    The top two rows are the input signal and raw output (identical to the top two rows in Fig.~7 in the main article).  
    The third row shows a background noise recording from the optical microphone when no audio signal is played. 
    The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
	The fifth and sixth rows show the results of applying the cascaded-notch filter and Wiener filter respectively.
    The seventh shows the results of applying both the cascaded-notch and Wiener filter (identical to the third row in Fig.~7 in the main article).
    The eight row shows the results of applying the logMMSE estimator (identical to the fourth row in Fig.~7 in the main article). 
	}
\end{figure}

@@

An inituitive way to remove noise is by directly subtracting the background noise from the recorded spectrum. 
The third row in Fig.~\ref{fig:op_mic_results} shows the interferometer background noise when there is no input signal.
The fourth row shows the spectrum obtained after subtracting the background noise spectrum, i.e. the result of subtracting the third row from the second row. 
We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified.

@@

A high-pass filter smoothly attenuates frequencies below some cut-off frequency. 
Applying a high-pass filter, with a cut-off frequency around $150\,{\rm Hz}$, to the signal spectrum works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics. However, the mains harmonics above $100\,{\rm Hz}$ remain. Using a high-pass filter with a higher cut-off can be used to mitigate this issue. However, it makes the played-back signal unrecognisable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
Often in speech processing, the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information to the human ear~\cite{SubjectiveComparison}. However, applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.

@@

\begin{figure}
	\includegraphics[width=.45\textwidth]{figures/psd_butterworth_14_6.pdf}
	\caption{\label{fig:psd_noise}
Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
}
\end{figure}

@@

One method to remove mains noise and its harmonics is to use a sequence of notch filters centred at each of the frequency bins we want to remove. This sequence is known as a ``cascade'' of notches. Here, the notches are smooth in comparison to the na{\"i}ve zeroing of each frequency (which looks like a rectangular comb).

The impulse response describes the reaction of a system to an input signal as a function of time. 
If a system is linear and time-invarient, the output signals are completely determined for any input signal using the impulse response. 
Finite inpulse response (FIRs) filters have finite length (they equal zero outside of a finite range). 
By contrast an infinite impulse response (IIR) filter has infinite length due to feedback. 
Typically FIR filters outperform IIR filters as they are always stable. 
However, IIR filters normally require fewer coefficients which in turn speeds up signal processing in comparison to FIR filters.

In this work we opt for a IIR notch filter. 
We write the filter in the form of a Z-transform, in which a discrete-time signal is converted to the complex frequency domain $z=e^{jf}$. 
The IIR notch filter we use here is given by~\citep{10.5555/541204}, 
\begin{equation}
    \label{eqn:notch}
    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}\,,
\end{equation}
where $\alpha$ and $\beta$ are parameters that control the filter. 
In Eqn.~\ref{eqn:notch}, $w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
We choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal. The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
\begin{equation}
    \label{eqn:notch15}
    H(z) = \prod_{k=0}^{14} H_k(z).
\end{equation}

We use the built-in MATLAB filter design toolbox in this work~\cite{MATLAB}. 
The amplitude (magnitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row in Fig.~\ref{fig:op_mic_results}.
We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.

Although the notch filter removes much of the mains hum sound, the filtered recording is not intelligible. Qualitatively it sounds more like a drum than a human voice. This is due to loss of voice information under the filter.
To overcome this, we turn to more advanced techniques, starting with the Wiener filter. Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.

@@

If we let the Wiener filter be non-causal (i.e. we estimate the current signal based on both past \emph{and future} observations), then we can represent Eqn.~\ref{eqn:wiener-hopf} in the frequency domain as
\begin{equation}
\label{eqn:wiener}
    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}\,,
\end{equation}
where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively. Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high and attenuates the signal where the SNR is low. The causal Wiener filter (that only makes estimates based on past observations) is similar. More detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.

@@

Speech enhancement of noisy channels is a classic problem in signal processing. 
In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech. 
This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes. This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}. We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}) to the recorded signal.

The final row in Fig.~\ref{fig:op_mic_results} shows the results of the logMMSE estimator.
Again there is significant noise reduction, however the speech remains indistict.
The logMMSE results are also presented in Section V.C. in the main text.

>>>>

In Section~V in the main article, we use the interferometer as an optical microphone; audio signals are used to vibrate the interferometer mirrors, the changing interference pattern is recorded, and we aim to recover the original signal from the resulting timeseries data.
In this Supplementary Material, we include a selection of signal processing techniques applied to the data which may be considered for undergraduate laboratory examples. 
We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.

@@

\begin{figure}
	\begin{center}
    %% trim order is <left> <lower> <right> <upper>
    % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
	\includegraphics[width=.5\textwidth]
% trim={1cm 3cm 1cm 1cm}]
{figures/combined_results_full_melatos_labelled.pdf}
%notch_and_wiener_superplot_v2.pdf}
	\end{center}
	\caption{\label{fig:op_mic_results}
	%\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
    The top two rows are the input signal and raw output (identical to the top two rows in Fig.~7 in the main article).  
    The third row shows a background noise recording from the optical microphone when no audio signal is played. 
    The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
	The fifth and sixth rows show the results of applying the cascaded-notch filter and Wiener filter respectively.
    The seventh shows the results of applying both the cascaded-notch and Wiener filter (identical to the third row in Fig.~7 in the main article).
    The eighth row shows the results of applying the logMMSE estimator (identical to the fourth row in Fig.~7 in the main article). 
	}
\end{figure}

@@

An intuitive way to remove noise is by directly subtracting the background noise from the recorded spectrum. 
The third row in Fig.~\ref{fig:op_mic_results} shows the interferometer background noise when there is no input signal.
The fourth row shows the spectrum obtained after subtracting the background noise spectrum, i.e.\ the result of subtracting the third row from the second row. 
We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified.

@@

A high-pass filter smoothly attenuates frequencies below some cut-off frequency. 
Applying a high-pass filter, with a cut-off frequency around $150\,{\rm Hz}$, to the signal spectrum works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics. However, the mains harmonics above $100\,{\rm Hz}$ remain. Using a high-pass filter with a higher cut-off can be used to mitigate this issue. However, it makes the played-back signal unrecognizable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
Often in speech processing, the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information to the human ear~\cite{SubjectiveComparison}. However, applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.

@@

\begin{figure}
	\includegraphics[width=.45\textwidth]{figures/psd_butterworth_14_6.pdf}
	\caption{\label{fig:psd_noise}
Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
After filtering we see significant attenuation (at least 3~dB) of all frequencies outside the band, including around 100~dB near 0~Hz due to even stronger attenuation far from the band (this sharp drop-off at low frequencies would be more similar to the high-frequency (3~kHz) behaviour if shown against a logarithmic scale for frequency as is sometimes done). However, the filter has little change on the harmonics within the band.
}
\end{figure}

@@

One method to remove mains noise and its harmonics is to use a sequence of notch filters centered at each of the frequency bins we want to remove. This sequence is known as a ``cascade'' of notches. Here, the notches are smooth in comparison to the na{\"i}ve zeroing of each frequency (which looks like a rectangular comb).

The impulse response describes the reaction of a system to an input signal as a function of time. 
If a system is linear and time-invariant, the output signals are completely determined for any input signal using the impulse response. 
Finite impulse response (FIRs) filters have finite length (they equal zero outside of a finite range). 
In contrast, an infinite impulse response (IIR) filter has infinite length due to feedback. 
Typically FIR filters outperform IIR filters as they are always stable. 
However, IIR filters normally require fewer coefficients which in turn speeds up signal processing in comparison to FIR filters.

In this work, we opt for an IIR notch filter. 
We write the filter in the form of a Z-transform, in which a discrete-time signal is converted to the complex frequency domain $z=e^{jf}$. 
The IIR notch filter we use here is given by~\citep{10.5555/541204}, 
\begin{equation}
    \label{eqn:notch}
    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}\,,
\end{equation}
where $\alpha$ and $\beta$ are parameters that control the filter. 
In Eqn.~\ref{eqn:notch}, $w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centered on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
We choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal. The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
\begin{equation}
    \label{eqn:notch15}
    H(z) = \prod_{k=0}^{14} H_k(z).
\end{equation}

We use the built-in MATLAB filter design toolbox in this work~\cite{MATLAB}. 
The amplitude (magnitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row in Fig.~\ref{fig:op_mic_results}.
We see that the mains harmonics are significantly attenuated compared to the background subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.

Although the notch filter removes much of the mains hum sound, the filtered recording is not intelligible. Qualitatively it sounds more like a drum than a human voice. This is due to the loss of voice information under the filter.
To overcome this, we turn to more advanced techniques, starting with the Wiener filter. Instead of just passively filtering different frequencies, this statistical technique optimizes an estimate of the original signal.

@@

If we let the Wiener filter be non-causal (i.e. we estimate the current signal based on both past \emph{and future} observations), then we can represent Eqn.~\ref{eqn:wiener-hopf} in the frequency domain as
\begin{equation}
\label{eqn:wiener}
    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}\,,
\end{equation}
where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively. Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high and attenuates the signal where the SNR is low. The causal Wiener filter (that only makes estimates based on past observations) is similar. A more detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.

@@

Speech enhancement of noisy channels is a classic problem in signal processing. 
In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech. 
This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimizes the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes. This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}. We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}) to the recorded signal.

The final row in Fig.~\ref{fig:op_mic_results} shows the results of the logMMSE estimator.
Again there is significant noise reduction, however, the speech remains indistinct.
The logMMSE results are also presented in Section V.C. in the main text.

### 2021-04-29 paper/ifo-tableTopGWs.tex
interference pattern image slightly larger
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

Sound is a commonly-used analogy when explaining gravitational-wave science. 
Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot. Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
We use audio signals throughout this work to simulate gravitational-wave-like signals.

@@

\begin{figure}
 \begin{center}
  \includegraphics[height=0.3\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
 \end{center}
 \caption{\label{fig:interference_pattern}
The interference pattern produced by the table-top Michelson interferometer.
This image was taken with the webcam aligned off the beam axis. %, which may explain why the (predicted circular) pattern appears elliptical.
The pink star indicates the point where the intensity timeseries measurements were taken.
The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
}
\end{figure}

>>>>

Sound is a commonly used analogy when explaining gravitational-wave science. 
Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot. Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
We use audio signals throughout this work to simulate gravitational-wave-like signals.

@@

\begin{figure}
 \begin{center}
  \includegraphics[width=0.35\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
 \end{center}
 \caption{\label{fig:interference_pattern}
The interference pattern produced by the table-top Michelson interferometer.
This image was taken with the webcam aligned off the beam axis. %, which may explain why the (predicted circular) pattern appears elliptical.
The pink star indicates the point where the intensity timeseries measurements were taken.
The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
}
\end{figure}
